---
title: "Data Manipulation"
author: "Austin Hsin"
date: "2023-09-19"
output: html_document
---

```{r echo = FALSE, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Litters + Pups Import

Select function for Litters and Pups datasets

```{r Litters + Pups}
options(tibble.print_min = 3)

litters_data = read_csv("./data_import_examples/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data_import_examples/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)

select(litters_data, group, litter_number, gd0_weight, pups_born_alive)
select(litters_data, group:gd_of_birth)
select(litters_data, -pups_survive)
select(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
rename(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
select(litters_data, starts_with("gd"))
select(litters_data, litter_number, pups_survive, everything())
relocate(litters_data, litter_number, pups_survive)

select(pups_data, litter_number, sex, pd_ears)
```

## Filter

Filter function with litters and pups datasets

```{r}
filter(pups_data, sex == 1)
filter(pups_data, sex == 2, pd_walk < 11)
```

## Mutate

Mutate function with litters and pups datasets

```{r}
mutate(litters_data,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
)
mutate(pups_data, pivot_minus7 = pd_pivot - 7)
mutate(pups_data, pd_sum = pd_ears + pd_eyes + pd_pivot + pd_walk)
```

## Arrange and |>

Arrange and |> function with litters and pups datasets where the second chunk
is an example of how we learned these functions and the third chunk is the piping example

Fourth chunk is the learning assessment 

```{r}
head(arrange(litters_data, group, pups_born_alive), 10)

litters_data_raw = read_csv("./data_import_examples/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data_clean_names = janitor::clean_names(litters_data_raw)
litters_data_selected_cols = select(litters_data_clean_names, -pups_survive)
litters_data_with_vars = 
  mutate(
    litters_data_selected_cols, 
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group))
litters_data_with_vars_without_missing = 
  drop_na(litters_data_with_vars, wt_gain)
litters_data_with_vars_without_missing

litters_data = 
  read_csv("./data_import_examples/FAS_litters.csv", col_types = "ccddiiii") |> 
  janitor::clean_names() |> 
  select(-pups_survive) |> 
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) |> 
  drop_na(wt_gain)

litters_data

read_csv("./data_import_examples/FAS_pups.csv", col_types = "ciiiii") |>
  janitor::clean_names() |> 
  filter(sex == 1) |> 
  select(-pd_ears) |> 
  mutate(pd_pivot_gt7 = pd_pivot > 7)
```

