---
title: "Data Tidying"
author: "Austin Hsin"
date: "2023-09-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(haven)
```

## pivot_longer

pivot_longer function using pulse dataset in a step by step way to show functionality

```{r}
pulse_df = 
  haven::read_sas("./data_import_Examples/public_pulse_data.sas7bdat") |>
  janitor::clean_names()

pulse_df

pulse_tidy_data = 
  pivot_longer(
    pulse_df, 
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    values_to = "bdi")

pulse_tidy_data

pulse_tidy_data = 
  pivot_longer(
    pulse_df, 
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi")

pulse_tidy_data
```

## pivot_longer summary

Tidying pulse dataset in one chunk/step

```{r}
pulse_df = 
  haven::read_sas("./data_import_examples/public_pulse_data.sas7bdat") |>
  janitor::clean_names() |>
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") |>
  relocate(visit) |>
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit)) |>
  arrange(id, visit)

print(pulse_df, n = 12)
```

## pivot_longer learning assessment

Tidying litters data for learning assessment

```{r}
litters_data = read_csv("./data_import_examples/FAS_litters.csv", col_types = "ccddiiii") |>
  janitor::clean_names()|>
  select(litter_number, gd0_weight, gd18_weight)|>
pivot_longer(
  gd0_weight:gd18_weight,
  names_to = "gd",
  values_to = "weight") |>
mutate(gd = recode(gd, "gd0_weight" = 0, "gd18_weight" = 18))
litters_data
```

Notes: need to read file and set col_types, clean names, then select your variables (can learn shortcuts for variable selection later)
pivot_longer chooses which variables to change, changing the names/prefixes to one variable and separating the values to a different variable
mutate adjusts the variable you want based on your function(replace, factor, recode, etc) where you must do var = function(var, insert changes)

## pivot_wider

```{r}
analysis_result = tibble(
  group = c("treatment", "treatment", "placebo", "placebo"),
  time = c("pre", "post", "pre", "post"),
  mean = c(4, 8, 3.5, 4)
)

analysis_result
```